Fast sparse period estimation

### 2015-04-12 poster/ITR_Poster.tex
minor

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Noncoherent least squares estimators \\ of carrier phase and amplitude}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, \\ I. Vaughan L. Clarkson and Barry G. Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported under the Australian Government's Australian Space Research Program}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\calC}{{\mathcal C}}
\begin{block}{Received signal model}
Receive $L$ noisy $M$-ary phase-shifted-keyed ($M$-PSK) symbols of the form
\[
y_i = a_0 s_i + w_i, \qquad i = 1, \dots, L.
\]
\begin{itemize}
\item $a_0 = \rho_0 e^{j\theta_0}$ is the unknown carrier phase $\theta_0$ and amplitude $\rho_0$,
\item $s_1,\dots,s_L$ are transmitted $M$-PSK symbols,
\item $w_1,\dots,w_L \in \complex$ are i.i.d. circularly symmetric complex random variables representing noise.
\end{itemize}

Interested in estimating $\rho_0$ and $\theta_0$.
\end{block}

\begin{block}{Least squares estimator}
If all symbols $s_1,\dots,s_L$ are known
\[
\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*.
\]
More interested in the practical situation where symbols are not known,
\[
\hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L} \sum_{i = 1}^L \abs{ y_i - a s_i }^2.
\]
This estimator can be computed in $O(L\log L)$ operations.
\end{block}

\begin{block}{Theorem (Almost sure convergence)}

Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
%\begin{equation}\label{eq:RiandPhii}
\vspace{-0.15cm}
\[
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
\]
%\end{equation}
and define the continuous function
\[
G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\] 
If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
\begin{enumerate}
\item \vspace{-0.15cm} $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item \vspace{-0.2cm} $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$,
\end{enumerate}
where $\sfracpart{\cdot}$ takes its argument `modulo $\tfrac{2\pi}{M}$' into $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
%\[
%\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
%\]  
\end{block}

\begin{block}{Theorem (Asymptotic normality)}
Let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
\[
g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
\]
Put $\hat{\lambda}_L = \sfracpart{\hat{\theta} - \theta_0}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$. %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then 
The distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
\[
\left( \begin{array}{cc} 
H^{-2} A & 0 \\
0 & \rho_0^2 B
\end{array} \right)
\]
as $L \rightarrow \infty$, where
\vspace{-0.15cm}
\[
H = G(0) -  2 \sin\big(\tfrac{\pi}{M}\big) \sum_{k = 0}^{M-1} g\big(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}\big),
\]
\vspace{-0.15cm}
\[
A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
\]

@@

\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
		\caption{Phase mean square error for $4$-PSK (QPSK)}
		\label{fig:plotphaseM4}
\end{figure}

\vspace{0.79cm}
\begin{figure}
	\centering
		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
		\label{fig:plotampM4}
\end{figure}

>>>>

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Fast sparse period estimation}
\author{Robby.~G.~McKilliam, I.~Vaughan.~L.~Clarkson \\ and Barry.~G.~Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported under the Australian Government's Australian Space Research Program}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\calC}{{\mathcal C}}
\begin{block}{Received signal model}

{
\def\vertgap{2}
\def\ph{0.4}
\def\T{1.1}
\newcommand{\raxis}{\draw[->] (-0.25,0) -- (8,0) node[above] {$\reals$}; \draw (0,-0.06)-- node[below] {$0$} (0,0.06) }
\newcommand{\pulse}[1]{ \draw[->,>=latex] (#1,0) -- (#1,1) }
\newcommand{\pulsewithnode}[2]{ \draw[->,>=latex] (#1,0) -- node[right] {#2} (#1,1) }
\begin{figure}[t]
	\centering
\begin{tikzpicture}[font=\small,scale=3]
    %\draw[very thin,color=gray] (-0.1,-1.1) grid (3.9,3.9);
\begin{scope}
    \raxis;
    \foreach \k in {0,1,...,6} { \k, \pulse{\T*\k+\ph}; }
    \draw[<->] (\T*2+\ph,1.2) -- node[above] {$T_0$} (\T*3+\ph,1.2) ;
    \draw[<->] (0.0,1.2) -- node[above] {$\theta_0$} (\ph,1.2) ; 
%    \draw[->] (-0.25,-2*\vertgap) -- (8,-2*\vertgap) node[above] {$\reals$};
\end{scope}
\begin{scope}[yshift=-\vertgap cm]
  \raxis;
  \node[below] at (\T*1+\ph-0.3,0) {$y_1$}; \pulsewithnode{\T*1+\ph-0.3}{$s_1=1$}; \draw[->] (\T*1+\ph,1.2) -- node[above] {$w_1$} (\T*1+\ph-0.3,1.2) ; 
  \node[below] at (\T*3+\ph+0.5,0) {$y_2$}; \pulsewithnode{\T*3+\ph+0.5}{$s_2=3$}; \draw[->] (\T*3+\ph,1.2) -- node[above] {$w_2$} (\T*3+\ph+0.5,1.2) ;
  \node[below] at (\T*6+\ph-0.4,0) {$y_3$}; \pulsewithnode{\T*6+\ph-0.4}{$s_3=6$};\draw[->] (\T*6+\ph,1.2) -- node[above] {$w_3$} (\T*6+\ph-0.4,1.2) ;
\end{scope}
\end{tikzpicture} 
%		\includegraphics{figs/figures-1.mps}
		\caption{Sparse noisy observations of a periodic sequence.  The original sequence~\eqref{eq:periodicprocess} is shown on the top and sparse noisy observations according to~\eqref{eq:sigmodel} are shown on the bottom.  In this example the observed subset is represented by integers $s_1=1$, $s_2=3$ and $s_3=6$.}
		\label{fig_stat_model}
\end{figure}
}

@@

\begin{figure}[p] 
  \centering 
  \begin{tikzpicture} 
\begin{groupplot}[
    group style={ 
        group name=my plots,
        group size=1 by 2,
        %ylabels at=edge left,
        xlabels at=edge bottom,
        xticklabels at=edge bottom,
        vertical sep=0pt
    },
    ylabel style={at={(0.06,0.5)}},
    footnotesize,
    width=25cm,
    height=15cm,
    tickpos=left,
    %ytick align=outside,
    %xtick align=outside,
    xmode=log,
    ymode=log,
    xmin=11e-5,
    xmax=0.4
%    enlarge x limits=false 
]

\nextgroupplot[
ylabel={MSE},
ymin=2e-10
]

      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};

            \addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN30q1.5geom10};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN30q5.0geom10};
\addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN30geom10};
\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN30geom10}; 
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN30geom10};
%       \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN30geom10}; 
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};

%      \addplot[mark=none,dashed,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN30q2};
 %     \addplot[mark=none,color=black] table {code/data/PeriodogramCLTN200};
 %     \addplot[mark=none,color=black,dashed] table {code/data/NormalisedLLSCLTN200}; 
 %     \addplot[mark=*,only marks,color=black,mark options={solid,fill=black,scale=0.5}] table {code/data/PeriodogramN200};
 %     \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q3};
 %     \addplot[mark=o,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN200q10};
%      \addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/NormalisedSamplingLLSN200}; 
       \node[font=\footnotesize] at (axis cs:0.2,4e-10) {$N=30$};

\nextgroupplot[
xlabel={$\sigma^2$},
xlabel style={at={(0.5,0.1)}},
ylabel={MSE},
ymin=2e-15,
ymax=8e-1,
legend style={
  legend columns=1,
  draw=none,
  %fill=none,
  legend pos=north west,
  %at={(0.75,1)},
  cells={anchor=west},
  font=\footnotesize
}
%ytickten={-1,-3,-5,-7,-9,-11,-13}
]

\addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN1200q1.5geom10};
 \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN1200q5.0geom10};
       \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN1200geom10};
\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN1200geom10};
      %\addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN1200geom10};
       \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN1200geom10};
%       \addplot[color=gray,mark=x,only marks,mark options={solid,color=black,fill=black,scale=1.3}] table {../code/data/NormalisedSamplingLLSN1200geom10};
 %      \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN1100geom10};
      %\addplot[mark=x,only marks,color=black,mark options={solid,fill=black,scale=1}] table {code/data/QuantisedPeriodogramFFTN1100q3};
 %     \legend{Periodogram Theory, Least squares theory, Quantised periodogram, Periodogram, Least squares}
     %\legend{Periodogram theory, $q=\tfrac{3}{2}T_{\text{max}}$, $q=4T_{\text{max}}$, Periodogram, Least squares, SLS2-adj}
     \node[font=\footnotesize] at (axis cs:0.32,2e-15) [anchor=south east] {$N=1200$};
   %  \node[rotate=16,font=\footnotesize] at (axis cs:7e-4,1.5e-14) {$N=1200$};

\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Periodogram Theory,Periodogram}

\end{groupplot}

  \end{tikzpicture}  
  \caption{Mean square error versus noise variance $\sigma^2$.  Trials are generated so that only $10\%$ of periods are observed on average. }\label{plot:multipleN}
\end{figure}

\begin{figure}
  \centering 
  \begin{tikzpicture}
%    \selectcolormodel{gray} 
    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.055,0.42)}},xlabel style={at={(0.5,0.1)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/LeastSquaresBenchmark};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
      \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newBenchmark};
      %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
      \legend{Periodogram, Quantised periodogram $q=5$, Seperable line search}
   \end{axis}
  \end{tikzpicture}  
  \caption{Computation time in milliseconds versus data length $N$.}\label{plot:benchmark}
\end{figure}

% \begin{figure}
% 	\centering
% 		\includegraphics[width=\linewidth]{../code/data/posterplot-2.mps}
% 		\caption{Phase mean square error for $4$-PSK (QPSK)}
% 		\label{fig:plotphaseM4}
% \end{figure}

% \vspace{0.79cm}
% \begin{figure}
% 	\centering
% 		\includegraphics[width=\linewidth]{../code/data/posterplot-1.mps}
% 		\caption{Amplitude mean square error for $4$-PSK (QPSK)}
% 		\label{fig:plotampM4}
% \end{figure}
